\documentclass[letterpaper,11pt]{article}

\usepackage{amsmath}
\usepackage{amssymb}
\usepackage[hmargin=1.25in,vmargin=1in]{geometry}
\usepackage{booktabs}
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage{lmodern}
\usepackage{microtype}
\usepackage{pdflscape}
\usepackage{subcaption}

\title{Coursework 6: STAT 570}
\author{Philip Pham}
\date{\today}

\begin{document}
\maketitle
\begin{enumerate}
\item In this question, you will implement the algorithm you described in
  Question 4 of Exercises 5. The algorithm derived in 4(b) will now be
  implemented for the prostate cancer data. These data are available in the R
  package \texttt{lasso2} and are named \texttt{Prostate}. Take $Y$ as $\log$
  prostate specific antigen and $x$ as log cancer volume. Implement the blocked
  Gibbs sampling algorithm using the prior given in the first equation of the
  aforementioned question, with $m_0 = m_1 =0$, $v_{00} = v_{11} = 2$,
  $v_{01} = 0$, and $a=b=0$. Run two chains, one with starting values
  corresponding to the unbiased estimates of the parameters and one starting
  from a point randomly generated from the prior
  $\pi\left(\beta_0, \beta_1\right)$. Report:
  \begin{enumerate}
  \item Histogram representations of the univariate marginal distributions
    $p\left(\beta_0 \mid y\right)$, $p\left(\beta_1 \mid y\right)$ and
    $p\left(\sigma \mid y\right)$, and scatterplots of the bivariate marginal
    distributions $p\left(\beta_0, \beta_1 \mid y\right)$,
    $p\left(\beta_0,\sigma \mid y\right)$, and
    $p\left(\beta_1,\sigma \mid y\right)$.
  \item The posterior means, standard deviations and 10\%, 50\%, 90\% quantiles
    of $\beta_0$, $\beta_1$, and $\sigma$.
  \item $\mathbb{P}\left(\beta_1 > 0.5 \mid y\right)$
  \item Justify your choice of \emph{burn-in} period. For example, you may
    present the trace plots $\beta_0^{(t)}$, $\beta_1^{(t)}$,
    $\left(\log\sigma^{2}\right)^{(t)}$ versus $t$.
  \end{enumerate}
\item Consider the data in Table \ref{tab:p2_data} contain data on a typical
  reliability experiment and give the failure stresses (in GPa) of four samples
  of carbon fibers of lengths 1, 10, 20 and 50mm.

  \begin{table}
    \tiny
    \centering
    \input{../hw3/p2_data.tex}
    \caption{Failure stress data for four groups of fibers.}
    \label{tab:p2_data}
  \end{table}

  \begin{enumerate}
  \item Consider a Bayesian analysis with a Weibull likelihood and independent
    lognormal priors,
    $\eta \sim \operatorname{LogNormal}\left(\mu_\eta, \sigma_\eta\right)$,
    $\alpha \sim \operatorname{LogNormal}\left(\mu_\alpha,
      \sigma_\alpha\right)$. Choose $\mu_\eta$, $\sigma_\eta$ so that the prior
    probability that $\eta$ lies between 0.5 and 30 is 0.9, and $\mu_\alpha$,
    $\sigma_\alpha$ so that the prior probability that α lies between 1 and 4 is
    0.9.

    \begin{description}
    \item[Solution:]
      $\log\eta \sim \mathcal{N}\left(\mu_\eta, \sigma_\eta\right)$ by
      definition of the lognormal distribution. Since $\log$ is a monotonic
      transformation,
      \begin{align}
        0.9 = \mathbb{P}\left(1/2 \leq \eta \leq 30\right)
        &= \mathbb{P}\left(\log\frac{1}{2} \leq \log\eta \leq \log 30\right) \nonumber\\
        &=
          \mathbb{P}\left(
          \Phi^{-1}\left(0.05\right) \leq
          \frac{\log\eta - \mu_\eta}{\sigma_\eta} \leq
          \Phi^{-1}\left(0.95\right)
          \right),
          \label{eqn:p2_prior_prob}
      \end{align}
      where $\Phi$ is the cumulative distribution function of a standard normal.

      Equation \ref{eqn:p2_prior_prob} implies that
      \begin{align*}
        \frac{\log\left(1/2\right) - \mu_\eta}{\sigma_\eta}
        &= \Phi^{-1}\left(0.05\right) \\
        \frac{\log 30 - \mu_\eta}{\sigma_\eta}
        &= \Phi^{-1}\left(0.95\right).
      \end{align*}

      Solving, we have that
      \begin{align*}
        \sigma_\eta
        &=
          \frac{\log 30 - \log\frac{1}{2}}{\Phi^{-1}\left(0.95\right) - \Phi^{-1}\left(0.05\right)} 
          \approx 1.2446 \\
        \mu_\eta
        &= \log 30 - \sigma_\eta \Phi^{-1}\left(0.95\right) =
          \log \frac{1}{2} - \sigma_\eta\Phi^{-1}\left(0.05\right)
          \approx 1.3540
      \end{align*}

      Repeating the calculating for $\alpha$, we have
      $\mu_\alpha \approx 0.6931$ and $\sigma_\alpha \approx 0.4214$.

      Calculations can be found in
      \href{http://nbviewer.jupyter.org/github/ppham27/stat570/blob/master/hw6/failure\_stresses.ipynb}{\texttt{failure\_stresses.ipynb}}.
    \end{description}
  \item Run MCMC for summarizing the posterior
    $p\left(\eta, \alpha \mid y\right)$, and implement this algorithm for each
    of the groups in Table \ref{tab:p2_data}. Report the posterior medians and
    90\% credible intervals for $\eta$ and $\alpha$ and give histograms
    representations of the posterior margins for $\eta$ and $\alpha$, and a
    scatterplot representation of $p\left(\eta, \alpha \mid y\right)$.

    \begin{description}
    \item[Solution:] 
    \end{description}
  \end{enumerate}
\item The data in Table \ref{tab:p3_data}, taken from Wakefield et al. (1994),
  were collected following the administration of a single 30mg dose of the drug
  Cadralazine to a cardiac failure patient. The response $y_i$ represents the
  drug concentration at time $x_i$, $i = 1,\ldots,8$. The most straightforward
  model for these data is to assume
  \begin{table}
    \centering
    \input{p3_data.tex}
    \caption{Concentrations of the drug Cadralazine (in mg/liter, $y_i$) as a
      function of time (in hours, $x_i$), for $i = 1,\ldots,8$.}
    \label{tab:p3_data}
  \end{table}

  \begin{equation}
    \log y_i = \mu\left(\beta\right) + \epsilon_i =
    \log\left[
      \frac{D}{V}\exp\left(-k_e x_i\right)
    \right]
    + \epsilon_i
    \label{eqn:p3_model}
  \end{equation}
  where
  $\epsilon_i \mid \sigma^2 \sim_\mathrm{iid}
  \mathcal{N}\left(0,\sigma^2\right)$, $\beta = \left[V,k_e\right]$ and the dose
  is $D = 30$. The parameters are the volume of distribution $V > 0$ and the
  elimination rate $k_e$.
  \begin{enumerate}
  \item For this model obtain expressions for:
    \begin{enumerate}
    \item The log-likelihood function $L\left(\beta, \sigma^2\right)$.
    \item The score function $S\left(\beta, \sigma^2\right)$.
    \item The expected information matrix $I\left(\beta, \sigma^2\right)$.
    \end{enumerate}
  \item Obtain the MLE, and give an asymptotic 95\% confidence interval for each
    element of $\beta$.
  \item Plot the data, along with the fitted curve.
  \item Using residuals, examine the appropriateness of the assumptions of the
    above model. Does the model seem reasonable for these data?
  \item The clearance $Cl = V \times k_e$ and elimination half-life
    $x_{1/2} = \log 2/k_e$ are parameters of interest in this experiment. Find
    the MLEs of these parameters along with asymptotic 95\% confidence intervals.
  \end{enumerate}
\end{enumerate}
\end{document}
